\normalfont \normalsize \textsc{University of Stuttgart} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Hauptseminar:\\The Logjam Attack \\ \horrule{2pt} \\[0.5cm]

### 2017-06-16 Ausarbeitung/paper.tex
TLS, Logjam finished, small primes started

\usepackage[square]{natbib}

@@

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\section*{Abstract}
\begin{abstract}
\citep{Adrian:2015:IFS:2810103.2813707} developed a new attack called Logjam to break many cryptographic systems using the Diffie-Hellman key exchange mechanism. This review elucidates its technical details and summarize overall results of their paper.
\end{abstract}
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\section{Introduction}
If it comes to network security, the Diffie-Hellman key exchange is the standard wildly used in practice and declared to be safe. However, \citep{Adrian:2015:IFS:2810103.2813707} could break the security on many popular and browser certified websites. They demonstrated this so called Logjam attack on the Transfer Level Security protocol that supports Diffe-Hellman as key exchange. They also implemented an attack exploiting weak and misconfigured groups using the Pohling-Hellman algorithm.

The second result the authors present is an estimate on cost of computing logs from key sizes of 1024 bits. I will briefly summarize their assumptions for the estimate and the results.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contain all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},\alpha:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = \alpha(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists\beta:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):\beta(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}

\subsection{Solutions}
\paragraph{Key Exchange}
The solution the authors propose is to draw private and public keys $(D,E)$ from finite fields of sizes equal to prime numbers $\mathrm{GF}(q)\cong\mathbb{Z}/q\mathbb{Z}$, i.e.
\begin{equation}
D, E \in \mathrm{GF}(q) = \{x \mod q \mid x \in \mathbb{Z}\}.
\end{equation}
By choosing the private key as the logarithm of the public key and an arbitrary basis $\alpha \in \mathrm{GF}(q)$,
\begin{eqnarray}
D = \alpha^E\mod q,\\
E = \log_\alpha Y \mod q,\\
1 \leq D,E \leq q-1,
\end{eqnarray}
the requirements 3. and 4. are fulfilled. Calculating $D$ from $E$ has logarithmic time complexity in $q$. Inversely, calculating $E$ from $D$ has a complexity of $\sqrt{q}$. Therefore the cipher-cryptanalyst ratio is exponential. Note that this is not a proven bound. We use the best known algorithms for these problem to get this ratio. Better algorithms have not been found for decades, therefore this method is considered safe. Also, one need to find good primes $q$ in order to assure worst or near worst case log-computation time.

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $D_A$, $D_B$, $E_A$, $E_B$ with a $q$ of sufficient size and exchange their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{E_AE_B} \mod q\\
& = & (\alpha^{E_A})^{E_B} \mod q\\
& = & {D_A}^{E_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{E_B})^{E_A} \mod q\\
& = & {D_B}^{E_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them have now the same session key $K_{A,B}$.

The trap-door property of this mechanism allows us to use it for one-way authentication. A signature $s$ of some user $A$ is $D_A(s)$. No one can forge it, since $D_A$ is private. $A$ cannot deny his or her signature, since everyone can confirm $E_A(D_A(s)) = s$.

@@

The algorithm for NFS in general is technically difficult to understand. That means there are many mathematical objects we must deal with and keep in mind at the same time. For simplicity I neglected some details for a better reading, while the drawback is that I need to refer to the original paper from time to time.

\subsection{Precomputation}
The precomputation consists of three steps. In Polynomial Selection, we find an appropriate representation for $\mathrm{GF}(q)$. Using this representation, we find equations that relate certain numbers with smaller primes in $\mathrm{GF}(q)$ in the Sieving step. These equations form matrices that can be used to find relations that lead to the logs of the base $\alpha$ in the Linear Algebra step. But first, one assumption must be made on $\alpha$.

@@

First, we represent $\mathrm{GF}(q)$ as $\mathbb{Z}/q\mathbb{Z}$ where elements are identified with their least non-negative residues. Then, choose and integer $m$ and polynomial $f:\mathbb{Z}\rightarrow\mathbb{Z}$ of degree $k$ such that $f$ is monic, irreducible over $\mathbb{Q}$ and $f(m)\equiv 0 \mod q$. We can find such a polynomial $f$ by choosing $m$ of suitable size and find coefficients $a_1,...,a_k$ such that
\begin{eqnarray}
p = \sum_{i=0}^{k}a_im^i
\text{ .}
\end{eqnarray}
It is required that $q\nmid\Delta_f$, where $\Delta_f$ is the discriminant of $f$. If the choice of $f$ or $m$ leads to $q\mid\Delta_f$, then try another $m$ or alter $f$ slightly, for example add $m$ to $a_i$ and subtract $1$ from $a_{i+1}$.

@@

The actual factor base $\mathcal{B}$ for which we want to find the logarithms consists of two parts: $\mathcal{B}_\mathbb{Q}$ the set of rational primes and $\mathcal{B}_K$ the set of good prime ideals in $O_K$, where $p,\Vert p'\Vert_1\leq B,\ \forall p\in\mathcal{B}_\mathbb{Q},p'\in\mathcal{B}_K$. Note that $B$ is the same bound for our base $\alpha$ (from our primary problem definition: Find $E$ for $D=\alpha^E\mod q$). Also, let $\mathcal{B}'\subset\mathcal{B}_\mathbb{Q}$ denote the set of prime factors of $\alpha$.

@@

From here we can derive factorizations for primes in $\mathcal{B}_K$ with the fact (not proven here) that there is an unique ideal $\hat{s}\in\mathcal{B}_K$ lying over $s$ and dividing $c+d\gamma$. Let $v_{\hat{S}}(c,d) = v_s(c,d)$ for this ideal and be zero for other ideals in $\mathcal{B}_K$ of norm $s$. Then we have the following
\begin{eqnarray}
\label{eqnSieve1}
c+dm = \prod_{s\in\mathcal{B}_\mathbb{Q}}s^{w_s(c,d)}
\end{eqnarray}
and
\begin{eqnarray}
\label{eqnSieve2}
c+d\gamma = \prod_{\hat{s}\in\mathcal{B}_K}\hat{s}^{v_{\hat{s}}(c,d)}
\text{ .}
\end{eqnarray}
Repeats the steps described so far to find more pairs $(c,d)$ to build equations \ref{eqnSieve1} and \ref{eqnSieve2} until we have at least $|\mathcal{B}|$ of these equations. For each equation, build a matrix, where its rows are the $w_s$'s and $v_{\hat{s}}$'s, we call this set $\mathcal{A}$. We modify $\mathcal{A}$ by extracting only those columns where elements are in $\mathcal{B}-\mathcal{B}'$. This means, we are only interested in solutions for primes $s\in\mathcal{B}'$. $\mathcal{A}$ is now the input for the next step.

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $E$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq ET\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provide a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,v_2,...,v_{r_0}\mod q$ for each $q\leq ET\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > ET\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $D$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(ET)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(ET)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (E^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

@@

Note that improved versions of this algorithm exist, such as \citep{joux2003improvements}.

\subsection{Individual Log Calculation}
Now that we have our database ready, we can calculate a private key $E$ given $D$. The first step is to find a sort of a decomposition of $D$ by sampling random integers $l\in{1,..,q-1}$ until we have found an $l$ that satisfies
\begin{equation}
\alpha^lD\equiv p_1p_2\cdots p_t\mod q
\text{ ,}
\end{equation}
where each $p_i$ should be "small", for example $\leq q^\frac{1}{k}$. If the discrete logs $x_i$ of each $q_i$ were known, we could compute $E$ with
\begin{equation}
E = \left(\sum_{i=1}^{t}x_i\right)-l \mod q
\text{ .}
\end{equation}
Finding the log for a certain $p_i$ is a similar procedure as the precomputation, where we computed the logs of factors of $\alpha$. We need to find a suitable polynomial that can be used to express relations between $p_i$ and the factor base $\mathcal{B}'$ in form of linear equations. These equations can also be found by sieving through co-prime pairs $(c,d)$ and cancelling out factors not in $\mathcal{B}'$. This gives us the desired result where $p_i$ can be represented by factors in $\mathcal{B}'$, i.e.
\begin{equation}
p_i \equiv \prod_{s\in\mathcal{B}'}s^{z'_s} \mod q
\end{equation}
and thus
\begin{equation}
\log_{\alpha}p_i = \sum_{s\in\mathcal{B}'}z'_s\log_{\alpha}s \mod (q-1)
\text{ ,}
\end{equation}
where $z'_s$ is some term that come from intermediate calculation step that I neglected for simplicity. As described above, the private key $E$ can now be calculate from $\log_{\alpha}p_i$'s.

%----------------------------------------------------------------------------------------
%	The Logjam Attack
%----------------------------------------------------------------------------------------
\section{The Logjam Attack}

%----------------------------------------------------------------------------------------
%	Other Attacks
%----------------------------------------------------------------------------------------
\section{Other Attacks}

%----------------------------------------------------------------------------------------
%	Other Attacks
%----------------------------------------------------------------------------------------
\section{Pohling-Hellman Algorithm}

>>>>

\usepackage[square]{natbib}
\usepackage{graphicx}

@@

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\section*{Abstract}
\begin{abstract}
In the field of study of security for computer networks, \citep{Adrian:2015:IFS:2810103.2813707} developed a new attack called Logjam to break many cryptographic systems that use Diffie-Hellman key exchange mechanism, where TLS 1.2 was the current standard at that time to establish secure end-to-end connections. This work elucidates technical details of the background for Logjam, in particular Diffie-Hellman key exchange, Number Field Sieve algorithm and TLS handshake, and summarizes the overall results.
\end{abstract}
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\section{Introduction}
If it comes to network security, the Diffie-Hellman key exchange is the standard wildly used in practice and declared to be safe. However, \citep{Adrian:2015:IFS:2810103.2813707} could break the security on many popular and browser certified websites. They demonstrated this so called Logjam attack on the Transfer Level Security protocol that supports Diffe-Hellman as key exchange. They also implemented an attack exploiting weak and misconfigured groups using the Pohling-Hellman algorithm.

The second result the authors present is an estimate on cost of computing logs from key sizes of 1024 bits. I will briefly summarize their assumptions for the estimate and the results.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\label{sec:DHKeyExchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contain all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},\alpha:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = \alpha(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists\beta:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):\beta(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}

\subsection{Solutions}
\paragraph{Key Exchange}
The solution the authors propose is to draw private and public keys $(D,E)$ from finite fields $\mathrm{GF}(q)\cong\mathbb{Z}/q\mathbb{Z}$, i.e.
\begin{equation}
D, E \in \mathrm{GF}(q) = \{x \mod q \mid x \in \mathbb{Z}\}.
\end{equation}
By choosing the private key as the logarithm of the public key and an arbitrary basis $\alpha \in \mathrm{GF}(q)$,
\begin{eqnarray}
D = \alpha^E\mod q,\\
E = \log_\alpha Y \mod q,\\
1 \leq D,E \leq q-1,
\end{eqnarray}
the requirements 3. and 4. are fulfilled. Calculating $D$ from $E$ has logarithmic time complexity in $q$. Inversely, calculating $E$ from $D$ has a complexity of $\sqrt{q}$. Therefore the cipher-cryptanalyst ratio is exponential. Note that this is not a proven bound. We use the best known algorithms for these problem to get this ratio. Better algorithms have not been found for decades, therefore this method is considered safe. Also, one need to find good primes $q$ in order to assure worst or near worst case log-computation time.

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $D_A$, $D_B$, $E_A$, $E_B$ from a number field $\mathrm{GF}(q)$ with a sufficient size $q$ and send each other their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{E_AE_B} \mod q\\
& = & (\alpha^{E_A})^{E_B} \mod q\\
& = & {D_A}^{E_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{E_B})^{E_A} \mod q\\
& = & {D_B}^{E_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them have now the same session key $K_{A,B}$.

Because of the trap-door property of this mechanism, it allows us to use it for one-way authentication. A signature $s$ of some user $A$ is $E_A(s)$. No one can forge it, since $E_A$ is private. $A$ cannot deny his or her signature, since everyone can confirm $D_A(E_A(s)) = s$.

@@

The algorithm for NFS in general is technically difficult to understand. This means that there are many mathematical objects we must deal with and keep in mind at the same time. For simplicity I neglected some details for a better reading, while the drawback is that I cannot show all intermediate steps and instead need to refer to the original paper from time to time.

\subsection{Precomputation}
\label{subsec:precomputation}
The precomputation consists of three steps. In Polynomial Selection, we find an appropriate polynomial representation for $\mathrm{GF}(q)$. Using this representation, we find equations that relate certain numbers with smaller primes in $\mathrm{GF}(q)$ in the Sieving step. These equations form matrices that can be used to find relations that lead to the logs of the base $\alpha$ in the Linear Algebra step. But first, one assumption must be made on $\alpha$.

@@

First, we represent $\mathrm{GF}(q)$ as $\mathbb{Z}/q\mathbb{Z}$ where elements are identified with their least non-negative residues. Then, choose an integer $m$ and polynomial $f:\mathbb{Z}\rightarrow\mathbb{Z}$ of degree $k$ such that $f$ is monic, irreducible over $\mathbb{Q}$ and $f(m)\equiv 0 \mod q$. We can find such a polynomial $f$ by choosing $m$ of suitable size and find coefficients $a_1,...,a_k$ such that
\begin{eqnarray}
q = \sum_{i=0}^{k}a_im^i
\text{ .}
\end{eqnarray}
It is required that $q\nmid\Delta_f$, where $\Delta_f$ is the discriminant of $f$. If the choice of $f$ or $m$ leads to $q\mid\Delta_f$, then try another $m$ or alter $f$ slightly, for example add $m$ to $a_i$ and subtract $1$ from $a_{i+1}$.

@@

The actual factor base $\mathcal{B}$ for which we want to find the logarithms consists of two parts: $\mathcal{B}_\mathbb{Q}$ the set of rational primes and $\mathcal{B}_K$ the set of good prime ideals in $O_K$, where all of these elements are bounded in size, i.e. $p,\Vert p'\Vert_1\leq B,\ \forall p\in\mathcal{B}_\mathbb{Q},p'\in\mathcal{B}_K$. Note that $B$ is the same bound for our base $\alpha$ (from our primary problem definition: Find $E$ for $D=\alpha^E\mod q$). Also, let $\mathcal{B}'\subset\mathcal{B}_\mathbb{Q}$ denote the set of prime factors of $\alpha$.

@@

From here we can derive factorizations for primes in $\mathcal{B}_K$ with the fact (not proven here) that there is an unique ideal $\hat{s}\in\mathcal{B}_K$ lying over $s$ and dividing $c+d\gamma$. Let $v_{\hat{S}}(c,d) = v_s(c,d)$ for this ideal and be zero for other ideals in $\mathcal{B}_K$ of norm $s$. Then we have the following
\begin{eqnarray}
\label{eqnSieve1}
c+dm = \prod_{s\in\mathcal{B}_\mathbb{Q}}s^{w_s(c,d)}
\end{eqnarray}
and
\begin{eqnarray}
\label{eqnSieve2}
c+d\gamma = \prod_{\hat{s}\in\mathcal{B}_K}\hat{s}^{v_{\hat{s}}(c,d)}
\text{ .}
\end{eqnarray}
Repeat the steps described so far to find more pairs $(c,d)$ to build equations \ref{eqnSieve1} and \ref{eqnSieve2} until we have at least $|\mathcal{B}|$ of these equations. For each equation, build a matrix, where its rows are the $w_s$'s and $v_{\hat{s}}$'s, we call this set $\mathcal{A}$. We modify $\mathcal{A}$ by extracting only those columns where elements are in $\mathcal{B}-\mathcal{B}'$. This means, we are only interested in solutions for primes $s\in\mathcal{B}'$. $\mathcal{A}$ is now the input for the next step.

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $E$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq ET\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provide a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows and columns that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,...,v_{r_0}\mod q$ for each $q\leq ET\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > ET\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $D$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(ET)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(ET)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (E^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

@@

\subsection{Individual Log Calculation}
\label{subsec:indiviualLogCalculation}
Now that we have our database ready, we can calculate a private key $E$ given $D$. The first step is to find a sort of a decomposition of $D$ by sampling random integers $l\in{1,..,q-1}$ until we have found an $l$ that satisfies
\begin{equation}
\alpha^lD\equiv p_1p_2\cdots p_t\mod q
\text{ ,}
\end{equation}
where each $p_i$ should be "small", for example $\leq q^\frac{1}{k}$. If the discrete logs $x_i$ of each $q_i$ were known, we could compute $E$ with
\begin{equation}
E = \left(\sum_{i=1}^{t}x_i\right)-l \mod q
\text{ .}
\end{equation}
Finding the log for a certain $p_i$ is a similar procedure as the precomputation, where we computed the logs of factors of $\alpha$. We need to find a suitable polynomial that can be used to express relations between $p_i$ and the factor base $\mathcal{B}'$ in form of linear equations. These equations can also be found by sieving through co-prime pairs $(c,d)$ and cancelling out factors not in $\mathcal{B}'$. This gives us the desired result where $p_i$ can be represented by factors in $\mathcal{B}'$, i.e.
\begin{equation}
p_i \equiv \prod_{s\in\mathcal{B}'}s^{z'_s} \mod q
\end{equation}
and thus
\begin{equation}
\log_{\alpha}p_i = \sum_{s\in\mathcal{B}'}z'_s\log_{\alpha}s \mod (q-1)
\text{ ,}
\end{equation}
where $z'_s$ is some term that come from intermediate calculation steps that I neglected for simplicity. As described above, the private key $E$ can now be calculate from $\log_{\alpha}p_i$'s.

Note that improved versions of the NFS algorithm for logarithm calculation exist, such as the one proposed in \citep{joux2003improvements}.
%----------------------------------------------------------------------------------------
%	Transport Layer Security
%----------------------------------------------------------------------------------------
\section{Transport Layer Security}
\label{sec:tls}
The document \citep{dierks2008transport} presents the full Transport Layer Security (TLS) protocol specification. I summarize the handshake section from there because the Logjam attack exploits the protocol flaw exactly here. An abbreviated message flow diagram is shown in Figure \ref{fig:tls}.
\begin{figure*}[h]
	\centering
	\includegraphics[width=10cm]{../Grafiken/TLS/tls}
	\caption{The message flow diagram of the TLS handshake, abbreviated.}
	\label{fig:tls}
\end{figure*}
\subsection{Hello Messages}
These messages initiate the handshake and exchange information about which cipher methods are supported on each side.
\paragraph{Hello Request} This message is optional and can be sent by the server at any time. The meaning is that the server wishes to renegotiate the connection. It expects to receive the \verb|ClientHello| message next.
\paragraph{Client Hello} With this message, the client wants to establish or re-establish a connection to the server. Several information are sent that are used for negotiation. The \verb|client_version| tells which TLS is used by the client. A random structure that consists of a time stamp and some random bytes are typically used for security checks, for example avoiding replay attacks. The client also offers a list of cipher and compression methods the server can choose from. It can use a session id to reuse security agreements from earlier or other session to abbreviate the handshake.
\paragraph{Server Hello} The server replies symmetrically by sending an \verb|server_version| which is lower or equals the \verb|client_version| and also a random structure that must not depend on the client's random structure. It selects the algorithms for cipher and compression. If none of these algorithms match, the handshake fails. If the client wished to reuse a session, the server sends the same session id back, otherwise generates a new one.
\paragraph{Server Hello Done}
This message indicates the end of a \verb|ServerHello| message, because its parameters does not have fixed length. The client should check all these parameter and also the server's certificate.

\subsection{Certification}
\paragraph{Certificate Request}
The server can request a certificate from the client if this is provided from the security method used. There are specific certificate types for the Diffie-Hellman key exchange method. The client needs to sign its certificate and make sure that the type matches the required one.
\paragraph{Client Certificate}
When the Server sent its \verb|ServerHelloDone| message and requests a certificate, the client sends this message. When there are no suitable certificates available, the content of this message should be empty. Then, the server decides to continue or to deny the handshake.
\paragraph{Certificate Verification}
When the server receives the certificate from the client, which turns out to be valid, the server sends this message to confirm the acknowledgement.

\subsection{Key Exchange}
TLS supports RSA and DH encryption. After the client has sent its certificate or, if it was necessary, after \verb|ServerHelloDone| was received, the client sends its public key to the server. The procedure for exchanging the session key basically goes according to Section \ref{sec:DHKeyExchange}.

\subsection{Finished}
These messages are the last ones in the handshake phase and the first ones that are encrypted with the newly exchanged session key. One participant sends a some encrypted content and the other answers with a proof that it can read the content, thus the handshake was successful. Now application data can be send supposedly safe.

%----------------------------------------------------------------------------------------
%	The Logjam Attack
%----------------------------------------------------------------------------------------
\section{The Logjam Attack}
The previous Sections explained the background for the Logjam attack more in detail than in the original work \citep{Adrian:2015:IFS:2810103.2813707}. Now I summarize the attack in short to see how everything comes together.

For preparation, use NFS precomputation to build log databases for some commonly used 512 bit primes (Section \ref{subsec:precomputation}). When the handshake (Section \ref{sec:tls}) we want to attack begins, catch the \verb|ClientHello| message as man in the middle and rewrite the offered cipher suit such that it only contains all weak ($\leq$ 512 bit) DH methods supported by the server. Since the client expects the server to choose from a strong cipher suit, we need to rewrite the \verb|ServerHello| message and replace the fake cipher suit with one that is strong but methodically matches the fake one. When keys are exchanged (Section \ref{sec:DHKeyExchange}), the client cannot detect any problem because the structure of these messages are the same regardless of which prime field is used. In the last step of the handshake, it is checked via \verb|Finished| messages whether the communication partner knows the secret. The attacker computes the secret (Section \ref{subsec:indiviualLogCalculation}) and replies accordingly. The client now beliefs that the attacker is the server and thereby lost its security.

The success of this attack depends on whether the servers being attacked support number fields of small primes or not. During the experiment it turned out that actually many servers do support those and some of them are even certificated being secure.

%----------------------------------------------------------------------------------------
%	Pohling-Hellman Algorithm
%----------------------------------------------------------------------------------------
\section{Exploiting Use of short Exponents}
Some application designers optimize the computation speed of $\alpha^E$ by choosing small $E$'s. This this can be unsafe, if the the prime $q$ is not chosen safe, which means that the subgroup generated by $\alpha$ in $\mathrm{GF}(q)$ has small factors. The authors of the Logjam attack also implemented an attack proposed by \citep{van1996diffie} and performed it on servers that have this weakness. This algorithm combines the Pollard's lambda method with Pohlig-Hellman decomposition. In the following, these two methods are described, before presenting the attack.

\paragraph{Pollard lambda} This method is a randomized one that solves a special case of the log problem $D = \alpha^E\mod q$, where the solution $E$ is known to be in range $b<E<b+w$ for some $b,w\in \mathbb{N}$. Using some random function $f:\mathbb{N}\rightarrow R$, two sequences of numbers are generated:
\begin{equation}
\begin{array}{lcl}
T & = & (y_0',...,y_N'),\\
W & = & (y_0,...,y_M)
\text{ .}
\end{array}
\end{equation}
Sequence $T$ starts with the value $y_0' = g^{b+w}$ and generate its sequence by $y_{i+1}' = y_i'\cdot g^{f(y_i')}$. For $W$, $y_0 = D$ and $y_{i+1} = y_i\cdot g^{f(y_i)}$. The log distance between $y_0'$ and $y_N'$,
\begin{equation}
\log_\alpha(y_N') = \log_\alpha(y_0') + d_N'
\text{ ,}
\end{equation}
is
\begin{equation}
d_N' = \sum_{i=0}^{N'-1}f(y_i')\mod (q-1)
\text{ .}
\end{equation}
and for $W$ analogously. The $T$ sequence stops at some $N$, while the $W$ sequence stops if there is a hit at some point $M$ such that $y_M = y_N'$ or if $d_M$ exceeds $w+d_N'$. In the first case, the solution
\begin{equation}
E = b + w + d_N' - d_M \mod (q-1)
\end{equation}
is found. Otherwise this iteration failed and we need to restart again.
